Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-10-19 clustering_infinite_model.tex
started on Section 4.6. Added task for Markus.

In addition, since $\mu_{\alpha, \nu}(B_\Pcal(p)) = \xi_{\alpha,\nu} e^{\frac{y}{2}}$, with $\xi_{\alpha,\nu} = \frac{4\alpha\BL{\nu}}{(2\alpha - 1)\pi}$, and
\[
	\Delta_{\Pcal}(p) = \iint_{\Rcal^2} T_\Pcal(p,p_1,p_2) \eta_{p}(x_1,y_1)\eta_{p}(x_2,y_2) \, dx1 \, dx_2 \, dy_1 \, dy_2,
\]
we have that
\begin{equation}\label{eq:relation_T_vs_Delta_P}
	\Exp{T_\Pcal(p)} = \xi_{\alpha,\nu}^2 \frac{e^{y}}{2} \Delta_\Pcal(p).
\end{equation}

@@

Before we analyze local clustering in the infinite limit model, we first establish some results for its degree distribution. We define the probability distribution for the degree of node $p = (x,y)$ as
\begin{equation}
	\rho_{p}(k) = \Prob{D_\Pcal(p) = k} = \Prob{\Po(\mu_{\alpha,\nu}(\BallPo{p})) = k},
\end{equation}
where $\Po(\lambda)$ denotes a Poisson random variable with mean $\lambda$. Recall from equation \eqref{eq:mu_B_y_P} that
\[
	\mu_{\alpha,\nu}(B(p)) = \xi_{\alpha,\nu} e^{\frac{y}{2}},
\]
only depends on the $y$-coordinate of $p$. Therefore, we will often write $\rho_y(k)$ instead of $\rho_{p}(k)$.

Using the above equation and the transformation of variables $z = \xi_{\alpha,\nu}e^{\frac{y}{2}}$, we compute
\begin{align*}
	\int_0^\infty \rho(y,k) e^{-\alpha y} \, dy 
    &= \frac{1}{\Gamma(k+1)} \int_0^\infty \left(\xi_{\alpha,\nu} e^{\frac{y}{2}}\right)^k 
    	e^{-\xi_{\alpha,\nu} e^{\frac{y}{2}}} e^{-\alpha y} \, dy\\
    &= \frac{(\xi_{\alpha,\nu})^{2\alpha}}{\Gamma(k+1)} \int_0^\infty 
    	\left(\xi_{\alpha,\nu} e^{\frac{y}{2}}\right)^{k - 2\alpha} e^{-\xi_{\alpha,\nu} e^{\frac{y}{2}}}
        \, dy\\
    &= \frac{(\xi_{\alpha,\nu})^{2\alpha}}{\Gamma(k+1)} \int_{\xi_{\alpha,\nu}}^{\infty} 
    	z^{k-2\alpha} e^{-z} \, dz\\
    &= (\xi_{\alpha,\nu})^{2\alpha}\frac{\Gamma(k-2\alpha)}{\Gamma(k+1)}
    	\left(1 - \frac{\Gamma^{-}(k-2\alpha, \xi_{\alpha,\nu})}{\Gamma(k - 2\alpha)}\right),
        \numberthis \label{eq:degree_distribution_P}
\end{align*}
where we recall that $\Gamma$ denotes the Gamma-function and $\Gamma^{-1}$ the lower-incomplete Gamma-function.

Since $\Gamma^{-}(k-2\alpha, \xi_{\alpha,\nu})\Gamma(k - 2\alpha) \to 0$ \PvdH{Find reference.} and $\Gamma(k-t)/\Gamma(k+1) \sim k^{-(t+1)}$ as $k \to \infty$ we deduce that
\begin{equation}\label{eq:degree_distribution_P_asymptotics}
	\Prob{D_\Pcal = k} = \int_0^\infty \rho_{y}(k) e^{-\alpha y} \, dy \sim (\xi_{\alpha,\nu})^{2\alpha} k^{-(2\alpha + 1)}.
\end{equation}

@@

@@

We wish to compute

\[   
	c_\infty = \alpha \int_0^{\infty} \Delta_\Pcal(y) e^{-\alpha y} \left(1-e^{-\xi e^{y/2}} - \xi e^{y/2} 
	e^{-\xi e^{y_0/2}}\right) \dd y, 
\]

>>>>

In addition, since $\mu_{\alpha, \nu}(B_\Pcal(p)) = \xi_{\alpha,\nu} e^{\frac{y}{2}}$, with $\xi_{\alpha,\nu} = \frac{4\alpha\nu}{(2\alpha - 1)\pi}$, and
\[
	\Delta_{\Pcal}(p) = \iint_{\Rcal^2} T_\Pcal(p,p_1,p_2) \eta_{p}(x_1,y_1)\eta_{p}(x_2,y_2) \, dx1 \, dx_2 \, dy_1 \, dy_2,
\]
we have that
\begin{equation}\label{eq:relation_T_vs_Delta_P}
	\Exp{T_\Pcal(p)} = \xi_{\alpha,\nu}^2 \frac{e^{y}}{2} \Delta_\Pcal(p).
\end{equation}

@@

Before we analyze local clustering in the infinite limit model, we first establish some results for its degree distribution. We define the probability distribution for the degree of node $p = (x,y)$ as
\begin{equation}
	\rho(p,k) = \Prob{D_\Pcal(p) = k} = \Prob{\Po(\mu_{\alpha,\nu}(\BallPo{p})) = k},
\end{equation}
where $\Po(\lambda)$ denotes a Poisson random variable with mean $\lambda$. Recall from equation \eqref{eq:mu_ball_y_P} that
\[
	\mu_{\alpha,\nu}(\BallPo{p}) = \xi_{\alpha,\nu} e^{\frac{y}{2}},
\]
only depends on the $y$-coordinate of $p$. Therefore, we will often write $\rho(y,k)$ instead of $\rho(p,k)$.

Using the above equation and the transformation of variables $z = \xi_{\alpha,\nu}e^{\frac{y}{2}}$, we compute
\begin{align*}
	\int_0^\infty \rho(y,k) e^{-\alpha y} \, dy 
    &= \frac{1}{\Gamma(k+1)} \int_0^\infty \left(\xi_{\alpha,\nu} e^{\frac{y}{2}}\right)^k 
    	e^{-\xi_{\alpha,\nu} e^{\frac{y}{2}}} e^{-\alpha y} \, dy\\
    &= \frac{(\xi_{\alpha,\nu})^{2\alpha}}{\Gamma(k+1)} \int_0^\infty 
    	\left(\xi_{\alpha,\nu} e^{\frac{y}{2}}\right)^{k - 2\alpha} e^{-\xi_{\alpha,\nu} e^{\frac{y}{2}}}
        \, dy\\
    &= \frac{(\xi_{\alpha,\nu})^{2\alpha}}{\Gamma(k+1)} \int_{\xi_{\alpha,\nu}}^{\infty} 
    	z^{k - 1 -2\alpha} e^{-z} \, dz\\
    &= (\xi_{\alpha,\nu})^{2\alpha} \frac{\Gamma^+(k - 2\alpha, \xi_{\alpha,\nu})}{\Gamma(k+1)},
\end{align*}
where we recall that $\Gamma$ denotes the Gamma-function and $\Gamma^{+}$ the incomplete Gamma-function.
We therefore conclude that
\begin{equation}\label{eq:degree_distribution_P}
	\Prob{D_\Pcal = k} = (\xi_{\alpha,\nu})^{2\alpha} \frac{\Gamma^+(k - 2\alpha, \xi_{\alpha,\nu})}{\Gamma(k+1)}
\end{equation}

Since $\Gamma^{+}(k-2\alpha, \xi_{\alpha,\nu})/\Gamma(k - 2\alpha) \sim k^{-(t+1)}$ as $k \to \infty$ \PvdH{Find reference.} we deduce that
\begin{equation}\label{eq:degree_distribution_P_asymptotics}
	\Prob{D_\Pcal = k} \sim (\xi_{\alpha,\nu})^{2\alpha} k^{-(2\alpha + 1)}
	\quad \text{as } k \to \infty.
\end{equation}

@@

\PvdH{@Markus: Please check that your $\xi$ equals $\xi_{\alpha,\nu} := \frac{4\alpha \nu}{(2\alpha - 1)\pi}$ in all computations below. If not please adjust the computations.}

@@

Now that we have an expression for $\Delta_{\Pcal}(y)$ we can compute the limit local clustering coefficient $c_\infty$ and the limit local clustering function $c_\infty(k)$, i.e. prove Theorem~\ref{thm:exact_expression_c_infty} and Theorem~\ref{thm:exact_expression_local_clustering}.

First we recall that
\begin{align*}
p_k := \Pee( D=k ) = \int_0^\infty e^{-\xi e^{y/2}} \frac{(\xi e^{y/2})^k}{k!}\alpha e^{-\alpha y} dy = 2\alpha \int_0^1 e^{-\frac{\xi}{z}} z^{2\alpha -k-1} dz =\frac{2\alpha \xi^{2\alpha}}{k!} \Gamma(k-2\alpha,\xi)
\end{align*}
\PvdH{@Markus: This result is taken from your notes. It seems to disagree with~\eqref{eq:degree_distribution_P}. Could you please check the other computations.}

We wish to compute
\[   
	c_\infty = \alpha \int_0^{\infty} \Delta_\Pcal(y) e^{-\alpha y} \left(1-e^{-\xi e^{y/2}} - \xi e^{y/2} 
	e^{-\xi e^{y_0/2}}\right) \dd y, 
\]
